---
title: "Assignment 7"
author: "Rodal Issema"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 6
    self_contained: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

pacman::p_load(tidyverse, haven, curl, magrittr, janitor, gt, captioner)

table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")

```


# Get, subset, and clean data

Used year of birth, biological sex, ever had sex, and race variables as variables in subsetted data

```{r}
#temporary destination folder
mytempdir <- tempdir()

# download and unzip the larger data set
myUrl <- "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip"

# zip file in $temp -- basename gets just the file name from the URL and not the URL path;
#   file.path stitches the tempdir() path to the file name
zipfile <- file.path(mytempdir, basename(myUrl))

# dta file in $temp
dtafile <- tools::file_path_sans_ext(zipfile)

# check if the dta file exists
if (!file.exists(dtafile)) {
    # if the dta file doesn't exist, check for the zip file
    # check if the zip file exists, download if necessary
    if (!file.exists(zipfile)) {
        curl::curl_download(url = myUrl, destfile = zipfile)
    }
    # unzip the downloaded zip file
    if (file.exists(zipfile)) {
        unzip(zipfile = zipfile, exdir = mytempdir)
    }
}

# if the data set has not been read, read it in
if (!exists("ah_dt")) {
    ah_dt <- haven::read_dta(dtafile)
}

# lowercase column names
colnames(ah_dt) %<>% str_to_lower()

#working with people born between 1977 to 1980
dt <- ah_dt %>% 
  mutate(race = case_when(h1gi6a == 1 | h1gi8 == 1 ~ 1,
                          h1gi6b == 1 | h1gi8 == 2 ~ 2,
                          h1gi6c == 1 | h1gi8 == 3 ~ 3,
                          h1gi6d == 1 | h1gi8 == 4 ~ 4,
                          h1gi6e == 1 | h1gi8 == 5 ~ 5,
                          TRUE ~ 9),
         race = factor(race, levels = c(1:5, 9), labels = c('White', 'Black/African American', 'American Indian', 'Asian/Pacific Islander', 'Other', 'Unknown/Missing'))) %>% 
  select(year_born = h1gi1y, sex = bio_sex, race, 	ever_have_sex = h1co1) %>%
  filter(year_born %in% 77:80) %>% 
  mutate(sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female")), 
         ever_have_sex = factor(ever_have_sex, levels = c(0, 1, 6, 8), labels = c("No", "Yes", "Refused", "DK")),
         year_born = factor(year_born, levels = 77:80, labels = 1977:1980))

#save df as RDS in temp folder
saveRDS(object = dt, file = file.path(mytempdir, "AHwave1_subset.RDS"))

```

# Tables

```{r}
# frequency table (counts and percentages)
count_pct <- function(x) {
  x %>%
    janitor::tabyl() %>% 
    adorn_totals("row") %>%
    adorn_pct_formatting() %>% 
    gt()
}

dt$year_born %>% count_pct()
```

*`r table_nums(name = "YearBorn_freq", caption = "Year of birth; frequency (%)")`*

```{r}
dt$sex %>% count_pct()
```
*`r table_nums(name = "Sex_freq", caption = "Biological sex; frequency (%)")`*


```{r}
dt$race %>% count_pct()
```

*`r table_nums(name = "Race_freq", caption = "Race; frequency (%)")`*

```{r}
dt$ever_have_sex %>% count_pct()
```

*`r table_nums(name = "EverHadSex_freq", caption = "Ever had sex; frequency (%)")`*


```{r}
# frequency table 2 variables, counts & percentages
dt %>% 
  group_by(year_born, ever_have_sex) %>% 
  tally() %>% 
  mutate(pct = round(100* n/sum(n), 2),
         n_pct = paste0(n, " (", pct, "%)")) %>% 
  select(-n, -pct) %>% 
  gt()

```

*`r table_nums(name = "YearBorn_EverHadSex", caption = "2x2 tabulation of year of birth and ever had sex variables")`*

# Figure
```{r}
dt %>% 
  filter(ever_have_sex %in% c("Yes", "No")) %>% 
  group_by(year_born, race) %>% 
  summarise(n = sum(ever_have_sex == "Yes")) %>% 
  group_by(year_born) %>% 
  mutate(pct = n/sum(n)) %>% 
  ungroup() %>% 
  ggplot(aes(year_born, pct,group = race, color = race) )+ 
  geom_line() +
  theme_minimal() +
  labs(x = "Year of birth", y = "Ever had sex (%)")
```

*`r figure_nums(name = "EverHadSex", caption = "Percentage of people who ever had sex by birth year and race")`*


# Rmd code
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
```

